Numerical error propagation in the HCP structural pre-processing pipelines

### 2018-03-15 paper.tex

We randomly selected 10 subjects from the HCP data release
S500. \todo{list the subjects.}

@@

We executed the pipelines using Docker
containers to simplify the deployment of different operating
system versions on execution platforms. The Docker images were built for the HCP
 pre-processing pipelines v3.19.0 (PreFreeSurfer and
FreeSurfer)~\cite{Glasser2013} in CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for reuse \todo{Add URL}. We
collected the provenance trace (tree of executed processes and files
accessed in read or write mode) for subject \todo{XXX} using system-call
interception as provided by the \reprozip tool~\cite{Chirigati2016}.

To support the computational challenge, we used the Canadian Brain
Imaging Research Platform (CBRAIN)~\cite{cbrain} to manage the
containers and data. CBRAIN is a Web platform developed at the
Montreal Neurological Institute (MNI) to tackle the computational
challenges faced by neuroimaging researchers.  Applications can be
ported to CBRAIN using the Docker and Singularity container
technologies through the Boutiques framework~\cite{boutiques}. Our
CBRAIN HCP plugins are publicly available at \url{github url} for
installation and re-use in CBRAIN instances.

@@

To further ensure that the observed errors are caused by operating system
updates (inter-os differences) and not by other factors, we
implemented the following sanity checks:
\begin{itemize}
\item The file checksums of every
subject are computed before and after processing to detect potential corruption during file transfers and other manipulations.
\item The checksum of the Docker container images is recorded with
  every execution to identify any unintentional package update. In
  addition, the list of packages with versions is printed by every
  task and checked for consistency.
\item Hardware information is captured to make sure that errors are
  not coming from different CPU architectures.
\item Every task is executed on its own copy of the data to ensure
  that input data is not modified by the pipeline. 
\end{itemize}

@@

We quantified the error in the image files using the Dice coefficient and normalized root mean square error (NRMSE) computed as follows:
\todo{put the equation, see Lalet's thesis}

@@

Once differences between conditions are quantified, we automatically
identify the steps in the pipeline responsible for such differences.
We use the \reprozip tool~\cite{reprozip} to record: (1) the tree of
processes executed by the pipeline and (2) the list of files read and
written by each process. This information is collected by system call
interceptions, through the \texttt{ptrace} system call \todo{is it
  POSIX or UNIX or Linux?}, and stored in a \texttt{SQLite}
database. The database stores information on all the processes which
are created by the \texttt{clone()} or \texttt{fork()} Linux
\todo{check if POSIX} system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
temporary files.

We reconstruct the tree of processes starting from the first process
created by the pipeline and defining its child processes as the ones
that were created through \texttt{clone()} or \texttt{fork()}. We
create a process graph from this tree by adding edges corresponding to
file dependencies between processes. A file dependency is defined
between processes A and B if a file written by A is read by
B. Figure~\ref{fig:graph-example} shows an example of a process graph
constructed at this stage \todo{Build the graph obtained on
  bin/example.sh and add it to the Figure}.
\begin{figure}
%  \includegraphics{brain\_classification}
  \caption{Process graph constructed from an example brain extraction script.}
  \label{fig:graph-example}
\end{figure}

A different process graph may be produced for every subject in every
execution condition. Among our dataset, we identified 4 types of
subjects with different numbers of T1w and T2w images. We verified
that the process graphs were identical for all subjects of the same
type, for all execution conditions. The remainder of the analysis is
done separately for each subject type. \todo{Check that}

@@

In a previous study~\cite{Scaria2017}, we showed that
pre-processing pipelines of the Human Connectome
Project~\cite{Glasser2013} were sensitive to operating system
variation (see Figure \ref{fig:1}).
\begin{figure}
%  \includegraphics{brain\_classification}
  \caption{Tissue classification produced by the HCP pre-processing
    pipelines on subject 105216 (CentOS 6 vs CentOS 7). \todo{Represent this in a different way.}}
  \label{fig:1}
\end{figure}

>>>>

We randomly selected 10 subjects from the HCP data release
S500. \todo{ list the subjects.}

@@

We executed the pipelines using Docker
containers to simplify the deployment of different operating
system versions on execution platforms. The Docker images were built for the HCP
 pre-processing pipelines v3.19.0 (PreFreeSurfer and
FreeSurfer)~\cite{Glasser2013} in CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. We
collected the provenance trace (tree of executed processes and files
accessed in read or write mode) for each type of subjects using system-call
interception as provided by the \reprozip tool~\cite{Chirigati2016}.

To support the computational challenge, we used the Canadian Brain
Imaging Research Platform (CBRAIN)~\cite{cbrain} to manage the
containers and data. CBRAIN is a Web platform developed at the
Montreal Neurological Institute (MNI) to tackle the computational
challenges faced by neuroimaging researchers.  Applications can be
ported to CBRAIN using the Docker and Singularity container
technologies through the Boutiques framework~\cite{boutiques}. Our
CBRAIN HCP plugins are publicly available at \url{https://github.com/glatard/cbrain-plugins-hcp} for
installation and re-use in CBRAIN instances.

@@

To further ensure that the observed errors are caused by operating system
updates (\emph{inter-os differences}) and not by other factors, we
implemented the following sanity checks:
\begin{itemize}
\item The file checksums of every
subject are computed before and after processing to detect potential corruption during file transfers and other manipulations.
\item The checksum of the Docker container images is recorded with
  every execution to identify any unintentional package update. In
  addition, the list of packages with versions is printed by every
  task and checked for consistency.
\item Hardware information is captured to make sure that errors are
  not coming from different CPU architectures.
\item Every task is executed on its own copy of the data to ensure
  that input data is not modified by the pipeline. 
\end{itemize}

@@

We quantified the error in the image files using the Dice coefficient and normalized root mean square error (NRMSE) computed as follows:
\begin{center}
\begin{equation}
RMSE = {\sqrt {\frac{1} {n}{\sum\limits_{t = 1}^n {(\hat{y}_{t} - {y}_{t} } })^{2} } }
\end{equation}
\end{center}

\begin{center}
\begin{equation}
NRMSE = {\frac{RMSE} {y_{max} - y_{min}}}
\end{equation}
\end{center}

@@

Once differences between conditions are quantified, we automatically
identify the steps in the pipeline responsible for such differences.
We use the \reprozip tool~\cite{reprozip} to record: (1) the tree of
processes executed by the pipeline and (2) the list of files read and
written by each process. This information is collected by system call
interceptions, through the \texttt{ptrace} as a UNIX base system call, and stored in a \texttt{SQLite}
database. The database stores information on all the processes which
are created by the \texttt{clone()} in Linux or \texttt{fork()} in UNIX-like
operating system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
temporary files.

We reconstruct the tree of processes starting from the first process
created by the pipeline and defining its child processes as the ones
that were created through \texttt{clone()} or \texttt{fork()}. We
create a process graph from this tree by adding edges corresponding to
file dependencies between processes. A file dependency is defined
between processes A and B if a file written by A is read by
B. Figure~\ref{fig:graph-example} shows an example of a process graph
constructed at this stage.

\begin{figure}
%  \includegraphics{brain\_classification}
  \includegraphics[scale=0.3]{images/simple_graph}
  \caption{Process graph constructed from an example brain extraction script.}
  \label{fig:graph-example}
\end{figure}

A different process graph may be produced for every subject in every
execution condition. Among our dataset, we identified 4 types of
subjects with different numbers of T1w and T2w images. We verified
that the process graphs were identical for all subjects of the same
type, for all execution conditions. The remainder of the analysis is
done separately for each subject type.

@@

In a previous study~\cite{Scaria2017}, we showed that
pre-processing pipelines of the Human Connectome
Project~\cite{Glasser2013} were sensitive to operating system
variation (see Figure \ref{fig:1}).
\begin{figure}
%  \includegraphics{brain\_classification}
  \includegraphics[width=0.5\textwidth, height=6cm]{images/brain_classification.png}
  \caption{Tissue classification produced by the HCP pre-processing
    pipelines on subject 105216 (CentOS 6 vs CentOS 7).}
  \label{fig:1}
\end{figure}
